### 2023-02-25 code/analysis_code/exploratory_analysis.qmd
Project Part 2

---
title: "Exploratory analysis "
author: "Kelly Hatfield"
date: "02/24/2023"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)

#Reminder of variable names
ls(mydata)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Histogram plots for the continuous outcomes.

First, Risk Standardized HAI Rate
```{r}
p1 <- mydata %>% ggplot(aes(x=HAI_039_RS_Rate)) + geom_histogram() 
plot(p1)
figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure_file, plot=p1)

```

Now Total Nurse Staffing per Resident per Day

```{r}
p2 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_histogram() 
plot(p2)
figure_file = here("results","Adj_Tot_Staff_prpd.png")
ggsave(filename = figure_file, plot=p2) 
```

Now showing relationship between staffing and HAI rates

```{r}
p3 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, y=HAI_039_RS_Rate, color=Staffing.Rating)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results","staff_hai.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by ownership type.

```{r}
p4 <- mydata %>% ggplot(aes(x=Ownership.Type, y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","staff_hai_ownership.png")
ggsave(filename = figure_file, plot=p4) 
```

# Notes

Might want to regroup ownership type into fewer categories (e.g., put all government and for-profit together.