### 2023-02-25 code/processing_code/processingfile_v1.qmd

---
title: "Cleaning Script"
author: "Kelly Hatfield"
date: "2023-02-23"
output: html_document
---

# Processing script

This code contains script to load and process/clean the data.

# Setup

Load needed packages. make sure they are installed.

```{r}

library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library (tidyverse) #for data processing/cleaning
library(readr)
```

# Data loading

```{r}
# path to data
# note the use of the here() package and not absolute paths
provider_location <- here::here("data","raw_data","NH_ProviderInfo_Jan2023.csv")
raw_provider <- read.csv(provider_location)

quality_location <- here::here("data","raw_data","CMS Nursing Home HAI039.csv")
raw_quality <- read_csv(quality_location)
View(raw_quality)
ls(raw_quality)

```

# Check data

Several ways of looking at the data. Lets start wtih a glimpse of the provider data.

```{r}
dplyr::glimpse(raw_provider)

```

# Cleaning

There are a lot of variables and observations in this dataset. We want to subset to just the variables of interest, first.

```{r}
raw_provider1 <- raw_provider[c("Federal.Provider.Number", "Average.Number.of.Residents.per.Day",
                   "Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day",
                   "Staffing.Rating", "Provider.State",   "Ownership.Type")]
skim(raw_provider1)

```

We see a spattering of missing values throughout the dataset with the majority of missing values for total staffing per resident per day. However for a dataset of over 15,000 it is mostly complete. We will leave in the missing values until we select which variables we will incorporate into our model.

Next, we need to explore the quality metric data.

```{r}
#Observe Variable Names
ls(raw_quality)
skim(raw_quality)
```

We know from our data gathering phase that nursing homes that didn't have enough eligible residents for measure calculation would not have a HAI measure calculated. We see 3,209 of those cases are missing here. We will exclude them from our analysis. Also, we can get rid of the unnecessary variables too.

```{r}

raw_quality1 <- raw_quality %>% 
  select (cms_certification_number_ccn, state, HAI_039_Observed, HAI_039_Obs_Rate, HAI_039_Denominator, HAI_039_RS_Rate)

skim(raw_quality1)
```

Now we have no more missing in the quality metrics!

Finally, we need to merge these datasets together. The common identifer variable is the CCN. But we will need to rename the variable from the provider table first. Then, get rid of the nursing homes with no HAI quality measures.

```{r}
raw_provider2 <- raw_provider1 %>% rename(cms_certification_number_ccn="Federal.Provider.Number")

raw_combined <- merge(raw_provider2, raw_quality1, by = "cms_certification_number_ccn")

raw_combined2 <- raw_combined %>% filter(!is.na(HAI_039_Observed))
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- raw_combined2
```

# Save data

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```