### 2023-04-21 code/analysis_code/analysis for paper.qmd
Peer Review Push 1

---
title: "Paper analysis "
author: "Kelly Hatfield"
date: "04/23/2023"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) 
library (tidyverse)

library(skimr)

library(ggplot2)

library(tidymodels)
library(performance)
library(stringr)
library(dplyr)

```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)

#Reminder of variable names
ls(mydata)
```

## Results

### Descriptive stats: staff and HAI rate

f2

figure2 = here("results","Staff_prpd_distribution.png")

ggsave(filename = figure2, plot=f2)

```{r}

mean(mydata$HAI_039_RS_Rate)

mytitlef1 <- "Figure: Distribution of adjusted rate of HAIs requring hospitalization by nursing home"

p1 <- mydata %>% ggplot(aes(x=HAI_039_RS_Rate)) + geom_histogram() + labs(
    x = "Adjusted rate of HAIs requring hospitalization",              # x axis title
    y = "Count of Nursing Homes",   # y axis title
    title=  str_wrap(mytitlef1, 50) ,      # main title of figure
    colour = "Continent"      # title of legend
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(face = "bold"))

figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure1, plot=p1)

mydata %>%
  group_by(COMP_PERF) %>% summarise( n = n(), percent=n()/sum(n()),
                                     mean = mean(HAI_039_RS_Rate))

vp <- ggplot(mydata, aes(x=COMP_PERF, y=HAI_039_RS_Rate)) + labs(x = " ", y = "Adjusted rate of HAIs requring hospitalization", title=  str_wrap(mytitlef1, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
vp

mytitlef2 <- "Figure: Distribution of total staff per resident per day by nursing home"

f2 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_histogram() + labs(y = "Count of Nursing Homes ", x = "Total Staffing per Resident per Day", title=  str_wrap(mytitlef2, 50))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(face = "bold"))

f2
figure2 = here("results","Staff_prpd_distribution.png")
ggsave(filename = figure2, plot=f2)

vp2 <- ggplot(mydata, aes(x=as.factor(Staffing.Rating), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + labs(x = " ", y = "Total Nurse Staffing Per Resident Per Day", title=  str_wrap(mytitlef1, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
vp2
figure2b = here("results","Staff_prpd_distribution_group.png")
ggsave(filename = figure2b, plot=vp2)

mydata %>%
  group_by(Staffing.Rating) %>% summarise( n = n(), percent=n()/sum(n()),
                                     mean = mean(Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day))

```

## Bivariate Analysis

```{r}
library(ggpubr)

mytitlef3 <- "Figure: Relationship of total staff per resident per day by nursing home with Adjusted HAI Rate by nursing home"
sp <- ggplot(mydata, aes(x= Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, y = HAI_039_RS_Rate)) + geom_point() + labs(y = "Adjusted HAI Rate ", x = "Total Nurse Staffing Per Resident Per Day", title=  str_wrap(mytitlef1, 75)) + geom_smooth(method=lm, se=FALSE)  + theme(plot.title = element_text(hjust = 0.5), panel.background = element_blank(), axis.line = element_line(color="black"), axis.line.x = element_line(color="black")) + stat_cor(method="pearson", label.x=0, label.y=20)

sp

figure3 = here("results","correlation.png")
ggsave(filename = figure3, plot=sp)

mydata %>%
  group_by(COMP_PERF) %>% summarise( n = n(), percent=n()/sum(n()),
                                     mean = mean(Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, na.rm=T))

mytitlef4a <- "Figure: Relationship of total staff per resident per day by nursing home with HAI Rating by nursing home"

vp4a <- ggplot(mydata, aes(x=as.factor(COMP_PERF), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + labs(x = " ", y = "Total Nurse Staffing Per Resident Per Day", title=  str_wrap(mytitlef4a, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
figure4 = here("results","staffingrate_HAIgroup.png")
ggsave(filename = figure4, plot=vp4a)

###### Save for later ######
mytitlef4b <- "Figure: Relationship of adjusted HAI rate with Staffing Rating by nursing home"
vp4b <- ggplot(mydata, aes(x=as.factor(Staffing.Rating), y=HAI_039_RS_Rate)) + labs(x = " ", y = "Adjusted HAI Rate", title=  str_wrap(mytitlef4b, 75))+ geom_violin(trim=FALSE) +  stat_summary(fun = "mean",
               geom = "point",
               color = "red")
vp4b

```

## Models

```{r}
# This enables the analysis to be reproducible when random numbers are used 
set.seed(555)

mydata$Staffing.Rating.F <- as.factor(mydata$Staffing.Rating)

# Put 3/4 of the data into the training set 
data_split <- initial_split(mydata, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)

recipe_2 <- recipe(HAI_039_RS_Rate ~ Staffing.Rating  + Ownership.Type + Average.Number.of.Residents.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data= train_data)

#model: linear regression using GLM engine
ln_mod <- linear_reg() %>% set_engine ("lm")

#workflow: tying together linear model and recipe of chocie
wflow2 <- workflow() %>% add_model(ln_mod) %>% add_recipe(recipe_2)

#Fitting the model to train dataset
model_fit2 <- wflow2 %>%  fit(data = train_data)

#Looking at model output;
model_fit2 %>% extract_fit_parsnip() %>%  tidy()

#Applying the model to unseen test data
predict(model_fit2, test_data)
model_aug2<- augment(model_fit2, test_data)

#pull RS HAI Rate and predictions
temps <- model_aug2 %>% pull(HAI_039_RS_Rate)
pred <- model_aug2 %>% pull(.pred)

#run rmse
rmse_vec(temps, pred)

#Looking at model output on all of the data
wflow2 %>%  fit(data = mydata) %>% extract_fit_parsnip() %>%  tidy()
```

# Appendix

## Table 1: Summary of Findings

Showing a bit of code to produce and save a summary table.

```{r}

#Appendix Tabe 1: Make a table of all the distributions of all variables included in the analysis univariate

summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)

```

##